---
title: "1-data_explore"
author: "Bernard"
date: "2020-04-17"
output: workflowr::wflow_html
editor_options: 
  chunk_output_type: console
---

## Introduction

### Import libraries

```{r, message=F, warning=F}

rm (list = ls())
source ("code/load_package.R")
source ("code/helper_func.R")

```

### Import data

```{r}

dat <- read_excel("data/Nack_T1_T5_long.xlsx")
var_names <- as.data.frame (read_excel("data/new_header.xlsx", col_names = FALSE))[,1]


names (dat)[1:length (var_names)] <- var_names
```


### Get column names

```{r, eval= FALSE}

names (dat) <- tolower(names (dat))

var_names <- names (dat)

```


### Remove redundant variables

```{r, eval= FALSE}
keep_var <- c ("time",
               "subj",
              "age",
              "sex",
              "grp",
              "vas_neck",
              "vas_arm",
              "vas_head",
              "ndi",
              "ses",
              "pcs",
              "tsk",
              "anx",
              "dep")

dat_sub <- dat[, names (dat) %in% keep_var]

dat_sub <- map_df (dat_sub, make_NA)

```

### Data exploration

#### Create report

```{r}

dat_sub %>%
  mutate (sex = factor (sex),
          grp = factor (grp)) %>%
  filter (time != 5) %>%
  pivot_wider(names_from = "time",
              names_sep = "_",
              values_from = c(vas_neck:dep)) %>%
  create_report(output_file = "explore_report.html",
                output_dir = "./output/")

```

#### Plot missing data

```{r fig.height=15}

dat_sub %>%
  filter (time != 5) %>%
  mutate (sex = factor (sex),
          grp = factor (grp)) %>%
  pivot_wider(names_from = "time",
              names_sep = "_",
              values_from = c(vas_neck:dep)) %>%
  plot_missing()

```

#### Dotplot

```{r, fig.height = 10}

df.plot <- dat_sub %>%
            filter (time != 5) %>%
            gather (-c(subj, age, grp, sex, time), key = var, value = val)

ggplot (df.plot ) +
  geom_point (aes(x = 1:nrow (df.plot), 
                  y = val, 
                  colour = as.factor (time), 
                  shape = as.factor (grp))) +
  facet_wrap(~ var, ncol = 5, scale = "free")


```

#### Descriptives for dataset

```{r, results="asis"}

tableby (time ~., data = dat_sub, digits = 2, digits.p = 2) %>%
  summary()

```

## Data preprocessing

### Restructure long to wide

Keep all variables except NDI in time = 2 and time = 3 as predictors
Keep NDI in time 4 as outcome

```{r, eval= FALSE}

target <- c("ndi")

predictor <- dat_sub %>%
  select (-target) %>%
  mutate (neckarm_pain = rowMeans(select (., matches ("neck|arm")), na.rm = TRUE)) %>%
  mutate (head_pain = vas_head) %>%
  select (-c(vas_neck, vas_arm, vas_head)) %>%
  gather (-c(subj, age, 
             sex, grp, time), key = var, value = val) %>%
  filter (time %in% c(2, 3)) %>%
  unite (var_time, c("var", "time")) %>%
  spread (key = var_time, value = val )


outcome <- dat_sub %>%
  gather (-c(subj, age, 
             sex, grp, time), key = var, value = val) %>%
  filter (time == 4) %>% # time == 4
  filter (var %in% target) %>%
  unite (var_time, c("var", "time")) %>%
  spread (key =var_time, value = val )


dat_wide <- outcome %>%
  inner_join(predictor, by = c("subj", "age",
                               "sex", "grp"))


```


### Check individuals variables with high missing

```{r, results="asis"}

missing_per_indv <- (rowSums(is.na(dat_wide))/ ncol (dat_wide[, -c(1:4)])) * 100 

dat_high_missing <- dat_wide[missing_per_indv == 100,] 

tableby ( ~., data = dat_high_missing, digits = 2, digits.p = 2) %>%
  summary(text = TRUE)

```


### Remove subjects with complete missing

```{r}
dat2 <- dat_wide[missing_per_indv < 100,] 
```

### Recode some variables

```{r}

dat2$sex <-  factor (dat2$sex) 
dat2$grp <-  factor (dat2$grp)  
dat2$subj <-  NULL

```

## Save data

```{r}

save (dat,
      dat2,
      file = "output/clean_dat.RData")

```


