---
title: "3-report"
author: "Bernard"
date: "2020-04-17"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction


## Import data

```{r}

rm (list = ls())

source ("code/load_package.R")

load ("output/bn_data.RData")
load ("output/clean_dat.RData")

imp.data = impute (fit, data = df.bn, method = "bayes-lw")

```

## Export results

### Correlation performance table

```{r}

var_order <- c("anx_2", "dep_2", "pcs_2", "ses_2", "tsk_2", "neckarm_pain_2", "head_pain_2",
               "anx_3", "dep_3", "pcs_3", "ses_3", "tsk_3", "neckarm_pain_3", "head_pain_3", "ndi_4")

var_label <- c("anx_3", "dep_3", "pcs_3", "ses_3", "tsk_3", "neckarm_pain_3", "head_pain_3",
               "anx_6", "dep_6", "pcs_6", "ses_6", "tsk_6", "neckarm_pain_6", "head_pain_6", "ndi_12")

corr.df_ord <- corr.df[var_order] 
correlation <- data.frame(Variable = names (corr.df_ord),
                          Value = corr.df_ord %>% round (2)) %>%
  mutate (Strength = ifelse (abs (Value) <= 0.3, "negligible",
                             ifelse (abs(Value) > 0.3 & abs(Value <= 0.5), "low",
                                     ifelse (abs(Value) > 0.5 & abs(Value <= 0.7), "moderate",
                                             ifelse (abs(Value) > 0.7 & abs(Value <= 0.9), "high",
                                                     "very high"))))) %>%
  mutate (Variable = factor (Variable, levels = var_order, labels = var_label))


ft <- flextable(correlation) %>%
      set_caption(paste0("Correlation between observed and predicted values")) %>%
      autofit()

my_path <- paste0("../manuscript/table_corr.docx")

my_doc <- read_docx()  %>% 
  body_add_flextable(ft)

print (my_doc, target = my_path)

```

### Descriptive plot

```{r}
df.plot <- dat2 %>%
  pivot_longer(cols = c("ndi_4" : "tsk_3"),
               names_to = "var",
               values_to = "val") %>%
  mutate (var = factor (var, levels = var_order, labels = var_label),
          grp = factor (grp, labels = c("general", "structured"))) %>%
  group_by(grp, var) %>%
  summarize (Mean = mean (val, na.rm = TRUE),
             Sd = sd (val, na.rm = TRUE))


f <- ggplot (df.plot) +
  geom_bar(aes (x = grp, y = Mean), colour = "black", fill = "black", stat = "identity") +
  geom_errorbar(aes (x = grp, ymin = Mean, ymax = Mean + Sd), width = 0.25) +
  facet_wrap(~ var, scales = "free") +
  theme_cowplot()

tiff(width = 10, height = 8, units = "in", res = 100, file = "../manuscript/fig1.tiff")
f
dev.off()

```

### Plot missing data of subset data

```{r, results="asis"}

dat2 <- dat2 %>%
  select (grp, sex, age, var_order)

names (dat2) [-c(1:3)] <- var_label

f <- plot_missing(dat2) +
  labs (y = "Percentage missing",
        x = "Variables")

tiff(width = 10, height = 8, units = "in", res = 100, file = "../manuscript/sm_fig1.tiff")
f
dev.off()

```

### BN network plot

```{r fig.height=10, fig.width=10}

rename_nodes <- function (net) {
  nodes (net)[grepl ("_3", nodes (net))] <- str_replace(nodes (net)[grepl ("_3", nodes (net))], "_3", "_6")
  nodes (net)[grepl ("_2", nodes (net))] <- str_replace(nodes (net)[grepl ("_2", nodes (net))] , "_2", "_3")
  nodes (net)[grepl ("_4", nodes (net))] <- str_replace(nodes (net)[grepl ("_4", nodes (net))] , "_4", "_12")

  return (net)
}


avg <- rename_nodes (avg)

demo.var <- c("grp", "sex")
early.var <-  grep ("_3", nodes (avg), value = TRUE)
late.var <-  grep ("_6", nodes (avg), value = TRUE)
out.var <-  grep ("_12", nodes (avg), value = TRUE)

# Rename columns
df.bn <- df.bn %>%
  select (grp, sex, var_order)

names (df.bn) [-c(1:2)] <- var_label

boot2 <- boot %>%
  map (rename_nodes)

bootstr = custom.strength(boot2, nodes = names(df.bn))
avg = averaged.network(bootstr, threshold = 0.7)
fit = bn.fit (avg, df.bn, method = "mle")

g = strength.plot(avg, 
                  bootstr, 
                  shape = "rectangle")

graph::nodeRenderInfo(g) = list(fontsize=18)
renderGraph(g)

arc_col <- data.frame(arcs = names (edgeRenderInfo(g)$col)) %>%
  separate(arcs, c("parent", "child"), sep = "~")

coef_fit <- coef(fit) 
coef_fit <- coef_fit[!map_lgl(coef_fit, is.matrix)]
coef_fit <- coef_fit[!map_lgl(coef_fit, is.table)]
coef_fit <-  coef_fit %>%
  unlist ()
coef_fit <- coef_fit[!grepl ("Intercept", names (coef_fit))]

coef_fit <- data.frame(arcs = names (coef_fit), coefs = coef_fit) %>%
  separate(arcs, c ("child", "parent"), sep = "[.]") 

new_col <- arc_col %>%
  left_join(coef_fit, by = c("parent", "child")) %>%
  mutate (coefs = replace_na(coefs,88)) %>%
  mutate (col = ifelse (coefs < 0, "red", 
                        ifelse (coefs == 88, "black", "blue"))) %>%
  mutate (col = ifelse (parent == "dep_6" & child == "ses_6", "red",
                        ifelse (parent == "ses_3" & child == "ses_6", "blue", col)))

new_arc_col <- new_col$col
names (new_arc_col) <-   names (edgeRenderInfo(g)$col)

nodeRenderInfo(g)$fill[demo.var] = "cornsilk"
nodeRenderInfo(g)$fill[early.var] = "tan1"
nodeRenderInfo(g)$fill[late.var] = "gold"
nodeRenderInfo(g)$fill[out.var] = "tomato"
edgeRenderInfo(g)$col <- new_arc_col
graph::nodeRenderInfo(g) = list(fontsize=14)


renderGraph(g)
tiff(width = 20, height = 10, units = "in", res = 100, file = "../manuscript/fig2.tiff")
renderGraph(g)
dev.off()


```


## Probing the expert system

### Check if `grp` is conditionally independent from `ndi_4`


```{r message=FALSE, warning=FALSE}

dsep(fit, x = "grp", y = "ndi_12")

```

### Evaluating the mediating infuence of `neckarm_pain_3` and `ses_6` on the `anx_3`-`ndi_12` relationship

#### Influence of `anx_3` and `ndi_12`

```{r fig.height=10, fig.width=10}

set.seed(123)

sim <- cpdist(fit, nodes = c("ndi_12", "anx_3"), n = 10^4,
               evidence = (TRUE))

m <- lm(ndi_12 ~ anx_3, data = sim)

coefs <- coef(m) 
b0 <- round(coefs[1], 2)
b1 <- round(coefs[2],2)
r2 <- round(summary(m)$r.squared, 2)
summary (m)

eqn <- bquote(italic(ndi_12) == .(b0) + .(b1)*italic(anx_3) * "," ~~ 
                  r^2 == .(r2))

tiff(width = 12, height = 8, units = "in", res = 100, file = "../manuscript/fig3.tiff")

 plot(sim$anx_3, sim$ndi_12, ylab = "12th mth NDI", xlab = "3rd mth Anxiety", col = "grey", cex.axis = 1.5, cex.lab = 1.5) + 
  abline(coef(m), lwd = 2) +
  abline(v = 0, col = 2, lty = 2, lwd = 2) +
  abline(h = 0, col = 2, lty = 2, lwd = 2) +
  text(x = 15, y = 80, labels = eqn, cex = 1.2)
 
dev.off()

```


#### Influence of `anx_3` and `ndi_12` when `ses_6` is constant

```{r echo=TRUE, message=FALSE, warning=FALSE}

set.seed(123)

avg.mutilated = mutilated(avg, evidence = list(ses_6 = 0))
strength.plot(avg.mutilated, bootstr)

fitted.mutilated = bn.fit (avg.mutilated , df.bn, method = "mle")
fitted.mutilated$ses_6 = list(coef = c("(Intercept)" = 0), sd = 0)

sim = cpdist(fitted.mutilated, nodes = c("ndi_12", "anx_3"), n = 10^4,
               evidence = (TRUE))

m <- lm(ndi_12 ~ anx_3, data = sim)

coefs <- coef(m) 
b0 <- round(coefs[1], 2)
b1 <- round(coefs[2],2)
r2 <- round(summary(m)$r.squared, 2)

summary (m)

eqn <- bquote(italic(ndi_12) == .(b0) + .(b1)*italic(anx_3) * "," ~~ 
                  r^2 == .(r2))


plot(sim$anx_3, sim$ndi_12, ylab = "12th mth NDI", xlab = "3rd mth Anxiety", col = "grey", cex.axis = 1.5, cex.lab = 1.5) + 
  abline(coef(m), lwd = 2) +
  abline(v = 0, col = 2, lty = 2, lwd = 2) +
  abline(h = 0, col = 2, lty = 2, lwd = 2) +
  text(x = 15, y = 100, labels = eqn, cex = 1.2)


```

#### Influence of `anx_3` and `ndi_12` when `neckarm_pain_3` is constant

```{r echo=TRUE, message=FALSE, warning=FALSE}

set.seed(123)

avg.mutilated = mutilated(avg, evidence = list(neckarm_pain_3 = 0))
strength.plot(avg.mutilated, bootstr)

fitted.mutilated = bn.fit (avg.mutilated , df.bn, method = "mle")
fitted.mutilated$neckarm_pain_3 = list(coef = c("(Intercept)" = 0), sd = 0)

sim = cpdist(fitted.mutilated, nodes = c("ndi_12", "anx_3"), n = 10^4,
               evidence = (TRUE))

m <- lm(ndi_12 ~ anx_3, data = sim)

coefs <- coef(m) 
b0 <- round(coefs[1], 2)
b1 <- round(coefs[2],2)
r2 <- round(summary(m)$r.squared, 2)

summary (m)

eqn <- bquote(italic(ndi_12) == .(b0) + .(b1)*italic(anx_3) * "," ~~ 
                  r^2 == .(r2))


plot(sim$anx_3, sim$ndi_12, ylab = "12th mth NDI", xlab = "3rd mth Anxiety", col = "grey", cex.axis = 1.5, cex.lab = 1.5) + 
  abline(coef(m), lwd = 2) +
  abline(v = 0, col = 2, lty = 2, lwd = 2) +
  abline(h = 0, col = 2, lty = 2, lwd = 2) +
  text(x = 15, y = 100, labels = eqn, cex = 1.2)


```

#### Influence of `anx_3` and `ndi_12` when `neckarm_pain_3` and `ses_6`is constant

```{r echo=TRUE, message=FALSE, warning=FALSE}

set.seed(123)

avg.mutilated = mutilated(avg, evidence = list(neckarm_pain_3 = 0, ses_6 = 0))
strength.plot(avg.mutilated, bootstr)

fitted.mutilated = bn.fit (avg.mutilated , df.bn, method = "mle")
fitted.mutilated$neckarm_pain_3 = list(coef = c("(Intercept)" = 0), sd = 0)
fitted.mutilated$ses_6 = list(coef = c("(Intercept)" = 0), sd = 0)

sim = cpdist(fitted.mutilated, nodes = c("ndi_12", "anx_3"), n = 10^4,
               evidence = (TRUE))

m <- lm(ndi_12 ~ anx_3, data = sim)

coefs <- coef(m) 
b0 <- round(coefs[1], 2)
b1 <- round(coefs[2],2)
r2 <- round(summary(m)$r.squared, 2)

summary (m)

eqn <- bquote(italic(ndi_12) == .(b0) + .(b1)*italic(anx_3) * "," ~~ 
                  r^2 == .(r2))


plot(sim$anx_3, sim$ndi_12, ylab = "12th mth NDI", xlab = "3rd mth Anxiety", col = "grey", cex.axis = 1.5, cex.lab = 1.5) + 
  abline(coef(m), lwd = 2) +
  abline(v = 0, col = 2, lty = 2, lwd = 2) +
  abline(h = 0, col = 2, lty = 2, lwd = 2) +
  text(x = 15, y = 100, labels = eqn, cex = 1.2)

```

### Evaluating the mediating infuence of `neckarm_pain_3` on the `head_pain_3`-`ndi_12` relationship

#### Influence of `head_pain_3` and `ndi_12`

```{r fig.height=10, fig.width=10}
set.seed(123)

sim <- cpdist(fit, nodes = c("ndi_12", "head_pain_3"), n = 10^4,
               evidence = (TRUE))

m <- lm(ndi_12 ~ head_pain_3, data = sim)

coefs <- coef(m) 
b0 <- round(coefs[1], 2)
b1 <- round(coefs[2],2)
r2 <- round(summary(m)$r.squared, 2)
summary (m)

eqn <- bquote(italic(ndi_12) == .(b0) + .(b1)*italic(head_pain_3) * "," ~~ 
                  r^2 == .(r2))

tiff(width = 10, height = 8, units = "in", res = 100, file = "../manuscript/fig4.tiff")

plot(sim$head_pain_3, sim$ndi_12, ylab = "12th mth NDI", xlab = "3rd mth headpain", col = "grey", cex.axis = 1.5, cex.lab = 1.5) + 
  abline(coef(m), lwd = 2) +
  abline(v = 0, col = 2, lty = 2, lwd = 2) +
  abline(h = 0, col = 2, lty = 2, lwd = 2) +
  text(x = 15, y = 75, labels = eqn, cex = 1.2)
 
dev.off()

```

#### Influence of `head_pain_3` and `ndi_12` when `neckarm_pain_3` is constant

```{r fig.height=10, fig.width=10}
set.seed(123)

avg.mutilated = mutilated(avg, evidence = list(neckarm_pain_3 = 0))
strength.plot(avg.mutilated, bootstr)

fitted.mutilated = bn.fit (avg.mutilated , df.bn, method = "mle")
fitted.mutilated$neckarm_pain_3 = list(coef = c("(Intercept)" = 0), sd = 0)

sim <- cpdist(fitted.mutilated, nodes = c("ndi_12", "head_pain_3"), n = 10^4,
               evidence = (TRUE))

m <- lm(ndi_12 ~ head_pain_3, data = sim)

coefs <- coef(m) 
b0 <- round(coefs[1], 2)
b1 <- round(coefs[2],2)
r2 <- round(summary(m)$r.squared, 2)
summary (m)

eqn <- bquote(italic(ndi_12) == .(b0) + .(b1)*italic(head_pain_3) * "," ~~ 
                  r^2 == .(r2))

plot(sim$head_pain_3, sim$ndi_12, ylab = "12th mth NDI", xlab = "2nd mth headpain", col = "grey", cex.axis = 1.5, cex.lab = 1.5) + 
  abline(coef(m), lwd = 2) +
  abline(v = 0, col = 2, lty = 2, lwd = 2) +
  abline(h = 0, col = 2, lty = 2, lwd = 2) +
  text(x = 15, y = 75, labels = eqn, cex = 1.2)
 

```
